A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-04 tex/codatadp.tex
more

\newcommand\jPrime{j\vthinspace'\negvthinspace}

@@

Datatypes are additionally characterized by an induction axiom schema:
%
\[
\begin{aligned}[t]
\text{Induction:}\quad
\AXC{$\forall x_1 \ldots \vthinspace x_{n_{ij}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{ij}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{ij}(\bar x)]$\quad\text{for each $i,j$}}
\UIC{$P_1[v_1] \mathrel\land \cdots \mathrel\land P_l[v_l]$}
\DP
\end{aligned}
\]
where the induction hypothesis $\mathit{IH}_{ij}^k(x)$
denotes either $P_{i'}(x)$ for some $i$ such that this is type
correct or else $\top$.
%
Ensures that we get the smallest model. E.g. for natural numbers, this rules out
nonstandard models, which could contain cyclic elements (e.g., n such that
n = Suc n) or even infinite acyclic elements $Suc (Suc \ldots)$.

>>>>

\newcommand\iPrime{i\vthinspace'\negvthinspace}
\newcommand\jPrime{j\vthinspace'\negvthinspace}

@@

Datatypes are additionally characterized by an induction axiom schema:
%
\[
\begin{aligned}[t]
\text{Induction:}\quad
\AXC{$\forall x_1 \ldots \vthinspace x_{n_{ij}}.\; \bigl(\bigwedge_{\,k\,} \mathit{IH}_{ij}^k[\bar x_k]\bigr) \longrightarrow P_i[C_{ij}(\bar x)]$\quad\text{for each $i,j$}}
\UIC{$P_1[v_1] \mathrel\land \cdots \mathrel\land P_l[v_l]$}
\DP
\end{aligned}
\]
where the induction hypothesis $\mathit{IH}_{ij}^k(x)$
denotes either $P_{\iPrime}(x)$ if there exists some $i$ such that this is type
correct or else $\top$.
%
The axiom schema ensures that we get a smallest model. For example,
for a datatype of natural numbers constructed from $\const{Z}$ and $\const{S}$,
this rules out nonstandard models, which could contain cyclic values (e.g.,
an $n$ such that $n \teq \const{S}(n)$) or even infinite acyclic values
$\const{S}(\const{S}(\ldots))$.
